DeepSphere: a graph-based spherical CNN

### 2019-09-24 main.tex
cosmo experiment

\todo{
\begin{itemize}
	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN.
	\item adjacency matrix: $A$ instead of $W$?
	\item anonymize DeepSphere citation
% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
	\item citet for inline, citep for parenthesis
	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method.
	\item PSD of climate data: how does it compare to cosmo and 3D objects?
	\item align the spheres in figure 1
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
	\item check footnotes
	\item climate data on spectrum plot
	\item result tables: only what supports the main message in the text, everything we have in the appendix
	\item Sampling, discretization, or pixelization of the sphere? Be consistent.
\end{itemize}
}
\newpage

@@

Another approach represents the discretized sphere as a graph connecting pixels according to the distance between them \citep{khasanova2017sphericalcnn, perraudin2019deepspherecosmo}.
While Laplacian-based graph convolutions are more efficient than spherical convolutions, they are not exactly equivariant \citep{defferrard2019deepsphereequiv}.
In this work, we argue that graph-based spherical CNNs strike an interesting balance, with a controllable tradeoff between cost and equivariance.
Experiments on multiple problems of practical interest show the competitiveness and flexibility of this approach.
% Furthermore, experiments show that, perhaps surprisingly and contrarily to previously published results \citep{cohen2019gauge} \todo{[Bronstein?]}, anisotropic filters might be an unnecessary price to pay.
% This paper focuses on improving the rotation equivariance of \citet{perraudin2019deepspherecosmo}, a scalable graph-based method designed for cosmological applications.

@@

\section{Is graph convolution equivariant to rotation? [2 pages]} \label{sec:equivariance}

@@

\todo{
Show: \\
* meet the desiderata \\
* DeepSphere V1 and V2 are equivalent in practice, wait on SHREC and cosmo exp \\
* anisotropy doesn't help \\
}

@@

The recognition of 3D shapes is a rotation invariant task: rotating an object doesn't change its nature.
While 3D shapes are usually represented as meshes or point clouds, representing them as spherical maps (\figref{depthmap}) naturally allows a rotation invariant treatment.
LIDAR actually acquires depth maps, which are a posteriori represented as point clouds.

@@

\subsection{Cosmological model discrimination} \label{sec:exp:cosmo}

\todo{
* other sphercial CNNs cannot scale to 10M pixels (tested on 10k at most) \\
* redo the experiment with the optimal graph \\
* DSv2 better than v1 -> equivariance has practical implications \\
* clear cost-accuracy tradeoff with n_neighbors \\
* show the baselines? (at least the 2D CNN?) \\
}

% \begin{table}[]
%     \centering
%     \begin{tabular}{c|c}
%          &  \\
%          &
%     \end{tabular}
%     \caption{Caption}
%     \label{tab:cosmo_newgraph}
% \end{table}

@@

To demonstrate the flexibility of modeling the discretized sphere by a graph, we collected historical measurements from weather stations scattered across the Earth.\footnote{\url{https://www.ncdc.noaa.gov/ghcn-daily-description}}
The spherical data is heavily non-uniformly sampled, with a much higher density of weather stations over North America than the Pacific (\figref{examples:ghcn:tmax}).
There are \todo{xx} stations in total.
For illustration, we devised two artificial tasks.
A dense regression that consists of predicting the temperature on a given day knowing the temperature on the 5 previous days.
A global regression that consists of predicting the day (represented as one period of a sine over the year) from temperature or precipitations.

>>>>

\todo{
\begin{itemize}
	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN.
	\item adjacency matrix: $A$ instead of $W$?
	\item anonymize DeepSphere citation
% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
	\item citet for inline, citep for parenthesis
	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method.
	\item PSD of climate data: how does it compare to cosmo and 3D objects?
	\item align the spheres in figure 1
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
	\item check footnotes
	\item climate data on spectrum plot
	\item result tables: only what supports the main message in the text, everything we have in the appendix
	\item Unify vocabulary. Sampling, discretization, or pixelization of the sphere? Signal, maps, functions.
\end{itemize}
}
\newpage

@@

Another approach represents the discretized sphere as a graph connecting pixels according to the distance between them \citep{bruna2013gnn, khasanova2017sphericalcnn, perraudin2019deepspherecosmo}.
While Laplacian-based graph convolutions are more efficient than spherical convolutions, they are not exactly equivariant \citep{defferrard2019deepsphereequiv}.
In this work, we argue that graph-based spherical CNNs strike an interesting balance, with a controllable tradeoff between cost and equivariance.
Experiments on multiple problems of practical interest show the competitiveness and flexibility of this approach.
% Furthermore, experiments show that, perhaps surprisingly and contrarily to previously published results \citep{cohen2019gauge} \todo{[Bronstein?]}, anisotropic filters might be an unnecessary price to pay.
% This paper focuses on improving the rotation equivariance of \citet{perraudin2019deepspherecosmo}, a scalable graph-based method designed for cosmological applications.

@@

\section{Graph convolution and equivariance [2 pages]} \label{sec:equivariance}
%\section{Is graph convolution equivariant to rotation? [2 pages]} \label{sec:equivariance}

@@

% Show:
% * meet the desiderata: scale
% * DeepSphere V1 and V2 are equivalent in practice: better equivariance doesn't translate to better performance
% * anisotropy doesn't help

@@

The recognition of 3D shapes is a rotation invariant task: rotating an object doesn't change its nature.
While 3D shapes are usually represented as meshes or point clouds, representing them as spherical maps (\figref{depthmap}) naturally allows a rotation invariant treatment.
LIDAR actually acquires depth maps, which are a posteriori represented as point clouds.
\todo{move to intro?}

@@

\subsection{Cosmological model classification} \label{sec:exp:cosmo}

Given observations, cosmologists estimate the posterior probability of cosmological parameters, such as the matter density $\Omega_m$ and the normalization of the matter power spectrum $\sigma_8$.
Those parameters are typically estimated by likelihood-free inference, which requires a function to predict the parameters from simulations.
% such as Approximate Bayesian computation (ABC)
As that is complicated to setup, prediction methods are typically benchmarked on the classification of spherical maps into cosmological model classes \citep{schmelze2017cosmologicalmodel}.
We used the same task, data, and setup as \citet{perraudin2019deepspherecosmo}: the classification of $720$ partial convergence maps ($1/12 \approx 8\%$ of the sphere) made of $10^6$ pixels from two $\Lambda$CDM cosmological models, ($\Omega_m = 0.31$, $\sigma_8 = 0.82)$ and ($\Omega_m = 0.26$, $\sigma_8 = 0.91)$, at a relative noise level of $4$.
% 60 maps * 12 parts = 720 parts
Convergence maps represent the distribution of over- and under-densities of mass in the universe \citep[see][for a review of gravitational lensing]{bartelman2010gravitationallensing}.

\begin{table}
    \centering
    \begin{tabular}{l c c c}
		\toprule
		& accuracy & conv. equiv. error & NN inv. error \\
		\midrule
		2D CNN baseline \citep{perraudin2019deepspherecosmo} & \todo{??} & \todo{can we?} \\
		8 neighbors, $\hat{t} = ??$, CNN variant, \citep{perraudin2019deepspherecosmo} & \todo{??} & \todo{??} \\
		8 neighbors, $\hat{t} = ??$, FCN variant, \citep{perraudin2019deepspherecosmo} & \todo{??} & \todo{??} \\
		8 neighbors, optimal kernel width $\hat{t} = ??$ & \todo{??} & \todo{??} \\
		20 neighbors, optimal kernel width $\hat{t} = ??$ & \todo{??} & \todo{??} \\
		40 neighbors, optimal kernel width $\hat{t} = ??$ & \todo{??} & \todo{??} \\
		\bottomrule
    \end{tabular}
	\caption{
		Results on the classification of partial convergence maps.
		\todo{While extreme non-equivariance (the 2D CNN baseline) penalizes performance, improving equivariance doesn't improve performance.}
	}
    \label{tab:cosmo}
\end{table}

\todo{
}

Results are shown in \tabref{cosmo}.
\todo{
	While extreme non-equivariance (the 2D CNN baseline) penalizes performance, better equivariance doesn't improve performance.
	It might be that the lack of built-in equivariance is compensated by learned equivariance (as done when augmenting data during training).}
\todo{
	Alternative: DSv2 better than v1 -> equivariance has practical implications.
	Clear cost-accuracy tradeoff with the number of neighbors.
}
This experiment moreover demonstrates DeepSphere's flexibility (using partial spherical maps) and scalability (competing spherical CNNs were tested on maps of at most $10,000$ pixels).

@@

To demonstrate the flexibility of modeling the discretized sphere by a graph, we collected historical measurements from about 10,000 weather stations scattered across the Earth.\footnote{\url{https://www.ncdc.noaa.gov/ghcn-daily-description}}
The spherical data is heavily non-uniformly sampled, with a much higher density of weather stations over North America than the Pacific (\figref{examples:ghcn:tmax}).
For illustration, we devised two artificial tasks.
A dense regression that consists of predicting the temperature on a given day knowing the temperature on the 5 previous days.
A global regression that consists of predicting the day (represented as one period of a sine over the year) from temperature or precipitations.
